Impact of Roe v. Wade’s Overturn: Abortion Restrictions and Infant Mortality in the U.S. Using a Difference-in-Differences Analysis

### 2024-11-29 paper/ExaminingtheimpactOverturnofRoev.Wade:BanningofAbortiononInfantMortalityRatesintheUnitedStatesUsingaDifference-in-DifferencesApproach.qmd
update paper with peer review

---
title: "Examining the impact Overturn of Roe v. Wade: Banning of Abortion on Infant Mortality Rates in the United States Using a Difference-in-Differences Approach"
subtitle: "The infant death rate increased by 0.285 deaths per 1,000 live births in states that became abortion-banned, compared to states where abortion remained legal."
author: 
  - Diana Shen
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion)."
date: today
date-format: long
abstract: "This study examines the impact of the Dobbs decision and subsequent abortion bans on infant mortality rates in the United States, focusing on data from 2021 to 2022. Using a Difference-in-Differences (DID) approach, we analyze trends in states where abortion remained legal compared to those where it became illegal after June 2022. Our results indicate that states enforcing abortion bans experienced an increase of 0.285 infant deaths per 1,000 live births, with compounded effects observed during the post-injunction period. Maternal age and race also emerged as significant predictors, with Black mothers facing disproportionately higher infant mortality rates. These findings underscore the compounded public health implications of restrictive abortion policies and the urgent need for targeted interventions to address racial and demographic disparities. By utilizing recent, high-quality data and rigorous modeling, this study provides essential understanding into the intersection of reproductive rights and infant health outcomes."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

library(tidyverse)
library(arrow)
library(knitr)
library(usmap)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(rstanarm)
library(fixest)  
library(caret)
```

@@

The overturning of Roe v. Wade on June 24, 2022, through the Supreme
Court’s decision in Dobbs v. Jackson Women’s Health Organization, ended
nearly five decades of constitutional protection for abortion rights.
This landmark ruling shifted authority over abortion regulations to
individual states, creating a patchwork of laws nationwide. The decision
reignited debates about reproductive rights and public health,
highlighting disparities in access to abortion services and their
effects on health outcomes, especially in states with restrictive
policies.

@@

The structure of this paper is as follows: @sec-data details the data sources and variables used in the analysis. @sec-model outlines the modeling approach, including assumptions and specifications of the Bayesian DID framework. In @sec-results, we present the findings, emphasizing key observations. In @sec-results, we present findings on the impact of abortion bans and the post-injunction period, highlighting their association with increased infant mortality. Key disparities by maternal age and race are discussed, along with model validation metrics like RMSE and $R^2$. Finally, @sec-discussion explores the implications of the results and suggests potential directions for future research. Additional details are provided in the appendices: [Appendix-@sec-data-details] elaborates on external data sources, [Appendix-@sec-model-details] provides an in-depth explanation of the model, and [Appendix-@sec-survey] introduces the idealized methodology for related qualitative analysis.

@@

```{r}
#| label: tbl-interact
#| tbl-cap: >
#|   Average Infant Death Rate by Abortion Legality and Injunction Timing: 
#|   States categorized as "Illegal Abortion and After Injunction" ('1') 
#|   exhibit a higher average infant death rate (per 1,000 live births) compared 
#|   to states in other categories, including "Legal Abortion and Before/After 
#|   Injunction" or "Illegal Abortion and Before Injunction" ('0'). 
#| echo: false
#| warning: false
#| message: false
analysis_data_state <- analysis_data|>mutate(illegal_after_injunction=as.numeric(as.character(abortion_illegal))*as.numeric(as.character(after_injunction)))|>
  group_by(illegal_after_injunction)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1))

@@

```{r}
#| label: tbl-summary
#| tbl-cap: >
#|   Demographic Distribution of Infant Mortality Cases by Year, Maternal Age, and Race (2021–2022): 
#|   The data highlights the proportional representation of cases across different years, 
#|   maternal age groups, and racial categories. Younger mothers and Black or African American 
#|   mothers show higher proportions of cases compared to other groups.
#| echo: false
#| warning: false
#| message: false
analysis_data|>mutate(year_of_death=as.factor(year_of_death))|>
  select(year_of_death,age_of_mother,mothers_single_race) |> 
  modelsummary::datasummary_skim()
```

@@

Addressing Public Health and Long-Term Implications of Restrictive Abortion Laws

@@

In @fig-ppcheckandposteriorvsprior-1 and @fig-ppcheckandposteriorvsprior-2 we implement a posterior predictive check for the first model and second model. It represents posterior predictive checks for first model and second model, illustrating how well each model captures the observed data. The posterior predictive checks for first model (policy-focused) and second model (policy and demographic factors) illustrate model fit. In @fig-ppcheckandposteriorvsprior-1, the observed ($y$) and replicated ($y_{rep}$) distributions align moderately, reflecting general trends but limited precision. @fig-ppcheckandposteriorvsprior-2 shows improved alignment, indicating that adding demographic factors enhances predictive accuracy and better captures the complexities of infant mortality rates.

In @fig-ppcheckandposteriorvsprior-3 and @fig-ppcheckandposteriorvsprior-4 we compare the posterior with the prior. The posterior vs. prior plots for first model and second model illustrate the role of data in informing the posterior estimates of the model parameters. In the first model @fig-ppcheckandposteriorvsprior-3, which focuses on policy-related predictors (e.g., abortion restrictions), the data significantly shift the posterior distributions for key variables, demonstrating their explanatory power. second model @fig-ppcheckandposteriorvsprior-4, incorporating additional demographic factors such as maternal age and race, shows a more refined and concentrated posterior distribution, emphasizing the added predictive utility of demographic variables. The comparison indicates that the second model provides better-informed parameter estimates, supporting the hypothesis that demographic factors play an essential role in understanding infant mortality in the context of restrictive abortion laws.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model for infant death rate fits, and is affected by, the data"
#| fig-subcap: ["First Model Posterior prediction check", "Second Model Posterior prediction check", "First Model Comparing the posterior with the prior", "Second Model Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")
pp_check(second_model) +
  theme_classic() +
  theme(legend.position = "bottom")
posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
posterior_vs_prior(second_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
prior_summary(first_model)
prior_summary(second_model)
```
The prior summaries for the two models, `first_model` and `second_model`, indicate that both use weakly informative priors. For the intercept, a normal prior with a mean of 0 and a standard deviation of 2.5 was specified, which adjusts to a scale of 12 after centering the predictors. The coefficients for both models are also assigned normal priors with a mean of 0 and a scale of 2.5, which are adjusted based on the predictors, resulting in scales such as [26.84, 28.43, 45.17] for the first model and [26.84, 28.43, 24.78,...] for the second model. The auxiliary parameter (sigma) follows an exponential prior with a rate of 1, adjusted to 0.2, while a decov prior regularizes the covariance structure in both models. Although the prior structure is similar, the second model incorporates more predictors, leading to a larger number of coefficients with slightly adjusted scales. These priors are designed to be weakly informative, allowing the data to primarily guide the posterior estimates while avoiding extreme parameter values.

## Distribution
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-area
#| layout-ncol: 2
#| fig-cap: "95% credibility intervals for fitted models"
#| fig-subcap: ["First Model", "Second Model"]

plot(
  first_model,
  "areas"
)
plot(
  second_model,
  "areas"
)
```
Credibility intervals are the Bayesian equivalent of confidence intervals, offering an understanding of the range of probable values for a parameter @tellingstories. Using Bayesian estimation, we derive a posterior distribution for each model coefficient. @fig-area-1 presents the 95% credibility intervals for coefficients in the first model, while @fig-area-2 illustrates the same for the second model. These intervals represent the range within which 95% of the posterior probability mass falls, offering a robust understanding of parameter uncertainty and the influence of predictors under the Bayesian framework @tellingstories.

@@

# Output the RMSE
print(paste("The RMSE is:", round(rmse_value, 2)))
```
The RMSE values from cross-validation indicate that the first model and second model perform similarly in predicting infant mortality rates, with RMSE values of 5.07 and 5.06, respectively. This similarity suggests that adding demographic factors such as maternal age and race in the second model has only a marginal impact on predictive accuracy compared to the first model, which primarily relies on policy-related variables like abortion legality and timing of the injunction. An average error of approximately 5 deaths per 1,000 live births indicates the magnitude of prediction errors, which may or may not be acceptable depending on the variability in the dataset. While the second model slightly reduces the error, its added complexity may not justify the trade-off if simplicity is a priority. However, if a nuanced understanding of demographic influences is essential, the second model provides a modest improvement in accuracy while offering more detailed insights into maternal and infant health disparities. The cross-validation process reinforces the reliability of these models by confirming their performance across multiple data splits, reducing the risk of overfitting.

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for the infant death rate model"
#| fig-subcap: ["First Model Trace plot", "First Model Rhat plot","Second Model Trace plot", "Second Model Rhat plot"]
#| layout-ncol: 2

@@

plot(first_model, "rhat")

plot(second_model, "trace")

@@

The trace plot in @fig-stanareyouokay-1 shows the sampling chains for the first model, indicating good mixing and convergence as the chains overlap and appear stable throughout. This suggests that the model has achieved convergence during sampling. The Rhat plot in @fig-stanareyouokay-2 demonstrates that all Rhat values are below 1.05, further confirming convergence and indicating that the model’s posterior samples are reliable. @fig-stanareyouokay-3 represents the trace plots for the second model, showing the posterior samples across four chains. The plots indicate good mixing, as the chains overlap well and stabilize around the same region, suggesting convergence.
@fig-stanareyouokay-4 is the Rhat diagnostic plot for the second model. All Rhat values are close to 1.00, confirming that the Markov chains have converged, and the posterior estimates are reliable. This supports the validity of the second model’s Bayesian estimation process.

@@

We draw on recommendations from @stantcheva2023 to minimize response biases. Common response biases identified in survey design include moderacy bias, extreme response bias, ordering bias, acquiescence bias, experimenter demand effect(EDE), and social desirability bias (SDB). Our survey primarily focuses on strategies to reduce moderacy bias, extreme response bias, ordering bias, and SDB. The detailed definitions of these biases are provided in [Appendix-@sec-definition].

@@

1.  What is your age?
    -   Under 18
    -   18–24
    -   25–34
    -   35–44
    -   45–54
    -   55+
    -   Prefer not to say
2.  What is your highest level of education?
    -   Less than high school
    -   High school graduate or equivalent
    -   Some college
    -   Bachelor’s degree
    -   Graduate or professional degree
    -   Prefer not to say

3. What is your marital status? - Single - Married - Divorced - Widowed -
Prefer not to say

@@

1.  Do you have any concerns or feedback regarding the survey, surveyor,
    or entity?
    -   Your feedback is important to us and will help ensure
        transparency and trust in the research process.

>>>>

---
title: "Impact of Roe v. Wade’s Overturn: Abortion Restrictions and Infant Mortality in the U.S. Using a Difference-in-Differences Analysis"
subtitle: "Infant death rates rose by 0.285 per 1,000 live births in abortion-restricted states compared to protected states."
author: 
  - Diana Shen
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion](https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion)."
date: today
date-format: long
abstract: "This study investigates the impact of the Dobbs decision and abortion bans on U.S. infant mortality rates using 2021–2022 data. A Difference-in-Differences (DID) analysis shows a 0.285 increase in infant deaths per 1,000 live births in states enforcing bans, with amplified effects post-injunction. Maternal age and race are key factors, with Black mothers experiencing higher mortality rates. These findings highlight the public health impact of restrictive abortion policies and the need for targeted interventions to address disparities. This study uses latest available data and rigorous modeling to explore the link between reproductive rights and infant health."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(arrow)
library(knitr)
library(usmap)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(rstanarm)
library(fixest)  
library(caret)
```

@@

The overturning of Roe v. Wade on June 24, 2022, through the Supreme
Court’s decision in Dobbs v. Jackson Women’s Health Organization, ended
nearly five decades of constitutional protection for abortion rights in
the United States. This landmark ruling shifted authority over abortion
regulations to individual states, creating a patchwork of laws
nationwide. The decision reignited debates about reproductive rights and
public health, highlighting disparities in access to abortion services
and their effects on health outcomes, especially in states with
restrictive policies.

@@

The structure of this paper is as follows: @sec-data details the data
sources and variables used in the analysis. @sec-model outlines the
modeling approach, including assumptions and specifications of the
Bayesian DID framework. In @sec-results, we present the findings,
emphasizing key observations. In @sec-results, we present findings on
the impact of abortion bans and the post-injunction period, highlighting
their association with increased infant mortality. Key disparities by
maternal age and race are discussed, along with model validation metrics
like RMSE and $R^2$. Finally, @sec-discussion explores the implications
of the results and suggests potential directions for future research.
Additional details are provided in the appendices: [Appendix-
@sec-data-details] elaborates on external data sources, [Appendix-
@sec-model-details] provides an in-depth explanation of the model, and
[Appendix- @sec-survey] introduces the idealized methodology for related
qualitative analysis.

@@

```{r}
#| label: tbl-interact
#| tbl-cap: >
#|   Average Infant Death Rate by Abortion Legality and Injunction Timing: 
#|   States categorized as "Illegal Abortion and After Injunction" ('1') 
#|   exhibit a higher average infant death rate (per 1,000 live births) compared 
#|   to states in other categories, including "Legal Abortion and Before/After 
#|   Injunction" or "Illegal Abortion and Before Injunction" ('0'). 
#| echo: false
#| warning: false
#| message: false
analysis_data_state <- analysis_data|>mutate(illegal_after_injunction=as.numeric(as.character(abortion_illegal))*as.numeric(as.character(after_injunction)))|>
  group_by(illegal_after_injunction)|>rename(`Abortion Restriction Post-Overturn`=illegal_after_injunction)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1))

@@

```{r}
#| label: tbl-summary
#| tbl-cap: >
#|   Demographic Distribution of Infant Mortality Cases by Year, Maternal Age, and Race (2021–2022): 
#|   The data highlights the proportional representation of cases across different years, 
#|   maternal age groups, and racial categories. Younger mothers and Black or African American 
#|   mothers show higher proportions of cases compared to other groups.
#| echo: false
#| warning: false
#| message: false
analysis_data|>mutate(year_of_death=as.factor(year_of_death))|>
  select(year_of_death,age_of_mother,mothers_single_race) |> 
  rename(`Year of Death` = year_of_death,
         `Age of Mother` = age_of_mother,
         `Mother's Single Race` = mothers_single_race)|>
  modelsummary::datasummary_skim()
```

@@

@@

In @fig-ppcheck-1 and @fig-ppcheck-2 we implement a posterior predictive
check for the first model and second model. It represents posterior
predictive checks for first model and second model, illustrating how
well each model captures the observed data. The posterior predictive
checks for first model (policy-focused) and second model (policy and
demographic factors) illustrate model fit. In @fig-ppcheck-1, the
observed ($y$) and replicated ($y_{rep}$) distributions align
moderately, reflecting general trends but limited precision.
@fig-ppcheck-2 shows improved alignment, indicating that adding
demographic factors enhances predictive accuracy and better captures the
complexities of infant mortality rates.

In @fig-posteriorvsprior1 and @fig-posteriorvsprior2 we compare the
posterior with the prior. The posterior vs. prior plots for first model
and second model illustrate the role of data in informing the posterior
estimates of the model parameters. In the first model
@fig-posteriorvsprior1, which focuses on policy-related predictors
(e.g., abortion restrictions), the data significantly shift the
posterior distributions for key variables, demonstrating their
explanatory power. second model @fig-posteriorvsprior2, incorporating
additional demographic factors such as maternal age and race, shows a
more refined and concentrated posterior distribution, emphasizing the
added predictive utility of demographic variables. The comparison
indicates that the second model provides better-informed parameter
estimates, supporting the hypothesis that demographic factors play an
essential role in understanding infant mortality in the context of
restrictive abortion laws.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| layout-ncol: 1
#| fig-cap: "Posterior Predictive Check: Examining how the model for infant death rate fits, and is affected by, the data"
#| fig-subcap: ["First Model Posterior prediction check", "Second Model Posterior prediction check" ]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")
pp_check(second_model) +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior1
#| fig-cap: "Comparing Prior and Posterior Distributions: Examining how the first model for infant death rate fits, and is affected by, the data"
posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior2
#| fig-cap: "Comparing Prior and Posterior Distributions: Examining how the second model for infant death rate fits, and is affected by, the data"
#| 
posterior_vs_prior(second_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| eval: false
#| echo: false
#| message: false
#| warning: false
#| include: false

prior_summary(first_model)
prior_summary(second_model)
```

## Prior Summary for Models

| **Model**        | **Component**     | **Specified Prior**                              | **Adjusted Prior**                               | **Description**                                         |
|--------------|--------------|--------------|-------------------|------------------|
| **First Model**  | Intercept         | Normal(0, 2.5)                                   | Normal(0, 12)                                    | Prior for intercept after centering predictors.         |
|                  | Coefficients      | Normal(\[0,0,0\], \[2.5,2.5,2.5\])               | Normal(\[0,0,0\], \[26.84,28.43,45.17\])         | Prior for coefficients, adjusted for predictor scaling. |
|                  | Auxiliary (sigma) | Exponential(1)                                   | Exponential(0.2)                                 | Prior for the standard deviation of residuals.          |
|                  | Covariance        | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Regularizing prior for covariance.                      |
| **Second Model** | Intercept         | Normal(0, 2.5)                                   | Normal(0, 12)                                    | Prior for intercept after centering predictors.         |
|                  | Coefficients      | Normal(\[0,0,0,...\], \[2.5,2.5,2.5,...\])       | Normal(\[0,0,0,...\], \[26.84,28.43,24.78,...\]) | Prior for coefficients, adjusted for predictor scaling. |
|                  | Auxiliary (sigma) | Exponential(1)                                   | Exponential(0.2)                                 | Prior for the standard deviation of residuals.          |
|                  | Covariance        | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Regularizing prior for covariance.                      |

: Prior Specifications for Bayesian Models {#tbl-priorsummary}

@tbl-priorsummary shows the prior summaries for the two models,
`first_model` and `second_model`, indicate that both use weakly
informative priors. For the intercept, a normal prior with a mean of 0
and a standard deviation of 2.5 was specified, which adjusts to a scale
of 12 after centering the predictors. The coefficients for both models
are also assigned normal priors with a mean of 0 and a scale of 2.5,
which are adjusted based on the predictors, resulting in scales such as
\[26.84, 28.43, 45.17\] for the first model and \[26.84, 28.43,
24.78,...\] for the second model. The auxiliary parameter (sigma)
follows an exponential prior with a rate of 1, adjusted to 0.2, while a
decov prior regularizes the covariance structure in both models.
Although the prior structure is similar, the second model incorporates
more predictors, leading to a larger number of coefficients with
slightly adjusted scales. These priors are designed to be weakly
informative, allowing the data to primarily guide the posterior
estimates while avoiding extreme parameter values.

## Distribution

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-area
#| layout-ncol: 1
#| fig-cap: "95% credibility intervals for fitted models"
#| fig-subcap: ["First Model", "Second Model"]

plot(
  first_model,
  "areas"
)

plot(
  second_model,
  "areas"
)
```

Credibility intervals are the Bayesian equivalent of confidence
intervals, offering an understanding of the range of probable values for
a parameter @tellingstories. Using Bayesian estimation, we derive a
posterior distribution for each model coefficient. @fig-area-1 presents
the 95% credibility intervals for coefficients in the first model, while
@fig-area-2 illustrates the same for the second model. These intervals
represent the range within which 95% of the posterior probability mass
falls, offering a robust understanding of parameter uncertainty and the
influence of predictors under the Bayesian framework @tellingstories.

@@

The RMSE values from cross-validation indicate that the first model and
second model perform similarly in predicting infant mortality rates,
with RMSE values of 5.07 and 5.06, respectively. This similarity
suggests that adding demographic factors such as maternal age and race
in the second model has only a marginal impact on predictive accuracy
compared to the first model, which primarily relies on policy-related
variables like abortion legality and timing of the injunction. An
average error of approximately 5 deaths per 1,000 live births indicates
the magnitude of prediction errors, which may or may not be acceptable
depending on the variability in the dataset. While the second model
slightly reduces the error, its added complexity may not justify the
trade-off if simplicity is a priority. However, if a nuanced
understanding of demographic influences is essential, the second model
provides a modest improvement in accuracy while offering more detailed
insights into maternal and infant health disparities. The
cross-validation process reinforces the reliability of these models by
confirming their performance across multiple data splits, reducing the
risk of overfitting.

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay1
#| fig-cap: "First Model: Checking the convergence of the MCMC algorithm for the infant death rate model"
#| fig-subcap: ["First Model Trace plot", "First Model Rhat plot"]
#| layout-ncol: 1

@@

plot(first_model, "rhat")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay2
#| fig-cap: "Second Model: Checking the convergence of the MCMC algorithm for the infant death rate model"
#| fig-subcap: ["Second Model Trace plot", "Second Model Rhat plot"]
#| layout-ncol: 1
plot(second_model, "trace")

@@

The trace plot in @fig-stanareyouokay1-1 shows the sampling chains for
the first model, indicating good mixing and convergence as the chains
overlap and appear stable throughout. This suggests that the model has
achieved convergence during sampling. The Rhat plot in
@fig-stanareyouokay1-2 demonstrates that all Rhat values are below 1.05,
further confirming convergence and indicating that the model’s posterior
samples are reliable. @fig-stanareyouokay2-1 represents the trace plots
for the second model, showing the posterior samples across four chains.
The plots indicate good mixing, as the chains overlap well and stabilize
around the same region, suggesting convergence. @fig-stanareyouokay2-2 is
the Rhat diagnostic plot for the second model. All Rhat values are close
to 1.00, confirming that the Markov chains have converged, and the
posterior estimates are reliable. This supports the validity of the
second model’s Bayesian estimation process.

@@

We draw on recommendations from @stantcheva2023 to minimize response
biases. Common response biases identified in survey design include
moderacy bias, extreme response bias, ordering bias, acquiescence bias,
experimenter demand effect(EDE), and social desirability bias (SDB). Our
survey primarily focuses on strategies to reduce moderacy bias, extreme
response bias, ordering bias, and SDB. The detailed definitions of these
biases are provided in [Appendix- @sec-definition].

@@

1.  What is your age?
    -   Under 18
    -   18–24
    -   25–34
    -   35–44
    -   45–54
    -   55+
    -   Prefer not to say
2.  What is your highest level of education?
    -   Less than high school
    -   High school graduate or equivalent
    -   Some college
    -   Bachelor’s degree
    -   Graduate or professional degree
    -   Prefer not to say
3.  What is your marital status? 
    - Single 
    - Married 
    - Divorced 
    - Widowed 
    - Prefer not to say

@@

1.  Do you have any concerns or feedback regarding the survey or entity?
    -   Your feedback is important to us and will help ensure
        transparency and trust in the research process.
